### 2019-09-22 papers/journal/2019_ACTA_Rebuttal_Rd2.txt

We would like to thank the reviewers for their very detailed and helpful comments. In the following, we elaborate on how we addressed the feedback in the revised version of our paper.

@@

% Reviewer 1:
  > My comments and concerns have been addressed in detail. 
    Thank you.
  Authors response: Thank you for your valuable comments.

% Reviewer 3:

> I believe that most issues highlighted previously have been 
  addressed adequately. In particular, the structure of the 
  revised manuscript has been improved significantly. 
  I still found the paper challenging to read, but this is 
  largely due to the complexity of the biological system and 
  the necessary encoding details presented in the manuscript.

  Authors response: Thank you for this feedback. As we are trying to reach an interdisciplinary audience we are forced to balance to some extent the biological detail with the technical discussions. We appreciate the feedback from the reviewers, which has helped us improve readability.

  > One point that I think requires further attention is the 
  motivation for using connectedness and the explanation of 
  this property, which is at the core of the proposed approach. 
  I did not find the additional explanations provided in the 
  response and the revised manuscript sufficient. Simply stating 
  that "The graph k-connectedness furnishes precisely" a test to 
  determine if information is missing from the model.

Authors response:

We have removed the word ``precisely'', which the Reviewer has correctly pointed out is too strong. We have replaced this by the sentence ``It has been shown that graph $k$-connectedness furnishes one such test''.

>  "If the data […] does not have a certain degree of connectivity,  this implies that some biological data has been missed" does not  justify connectedness as a fundamental and required biological  property that all valid models should respect. The intuition of  connectedness as a robustness property provided elsewhere in the  manuscript helps but it is still not clear what evidence supports  the fact that VTSs are indeed robust. In addition, the problem as  formulated does not seem to seek the most robust networks as  explained in the following.

Author response: We have added several paragraphs to motivate our focus on k-edge-connectedness:

``The informativeness and usefulness of using $k$-connectedness is motivated by the following example. ... In summary, the higher the edge-connectedness of a measured VTS graph, the more of the underlying biological constraints can be satisfied by appropriately placing SNAREs on each node and edge.''

  > Finally, it is not clear why the goal is to find the minimum graph connectivity, given that larger connectivity corresponds to a more robust network. I was expecting an approach involving incrementing k to find the smallest value that makes the formula unsatisfiable, thus identifying the maximum graph connectedness possible for a given VTS with all other constraints being satisfied. However, to the best of my understanding, this is not what the paper is currently proposing, or else further clarifications are needed.

Author response: We apologise for the confusing structure of our text. To clarify this point we have added a new paragraph:

``It is perhaps counter-intuitive that, beyond a certain degree of edge-connectedness, any VTS graph can be implemented as a biological system. This is because we have not placed any constraints on the number of distinct SNAREs and other trafficking molecules a cell contains. In practice there is an upper bound on the number of molecules a cell can genetically encode. This question (``What is the maximum edge-connectedness possible given a certain number of SNAREs?'') is complementary to the question we have asked here (``What is the minimum edge-connectedness possible given any number of SNARES operating under certain physical constraints?'').''

@@

  > I also found the formulation of the main problems with respect to connectedness somewhat confusing. For example, in Sec. 5.1.1 and Sec. 5.1.2 the problems are formulated as a search for VTS G such that G is not k-connected. Yet, Sec. 5.1 specifies that "In case the formula is satisfiable the procedure terminates and we report the value of k", which seems contradictory.

  Authors response: In Sec. 5 we try to give a high-level 
  explanation of the procedure, followed by a formal search statement from the perspective of an efficient implementation. 
  This difference was only hinted but not explicitly stated. We have  now rewritten the parts of the text to address the distinction between high-level overview and the implementation. We have also removed the specific search problems (as suggested by Reviewer #3 in the first revision) for further apparent accessibility of the subject matter.

### 2019-09-22 papers/journal/bio.tex
Mukund's push

The biological problem often boils down to such an analysis. A cell biologist might determine which molecules flow between which set of compartments, and biochemical experiments could be used to see how these molecules activate one another. We can then ask: is this a complete and consistent description? That is, do all the required properties listed above hold, given what the experimentalists have told us? It is often the case that biological data is missing. For example, only a few of the dozens of molecules involved in real VTSs have been mapped out. Therefore, it is extremely likely that the description provided by the cell biologist is incomplete. We can use our model to find out which properties have failed to hold, and thus prescribe new experiments in order to fill in the missing information.

Can we find a simple test to see whether any information is missing, given the experimental data? It has been shown that graph $k$-connectedness furnishes precisely such a test \cite{shukla2017discovering}.
%
If the data provided by cell-biologists, suitably represented as a graph, does not have a certain degree of connectivity, this implies that some biological data has been missed. (The converse is not true: even if the required connectivity does hold, there might of course be more information missing.)

>>>>

The biological problem often boils down to such an analysis. A cell biologist might determine which molecules flow between which set of compartments, and biochemical experiments could be used to see how these molecules activate one another. We can then ask: is this a complete and consistent description? That is, do all the required properties listed above hold, given what the experiments have told us? It is often the case that biological data is missing. For example, only a few of the dozens of molecules involved in real VTSs have been mapped out. Therefore, it is extremely likely that the description provided by the cell biologist is incomplete. We can use our model to find out which properties have failed to hold, and thus prescribe new experiments in order to fill in the missing information.

Can we find a simple test to see whether any information is missing, given the experimental data? It has been shown that graph $k$-connectedness furnishes one such test \cite{shukla2017discovering}.

The informativeness and usefulness of using $k$-connectedness is motivated by the following example. A necessary condition for the VTS graph to be in a steady state is that the graph is strongly connected. By Robbins' Theorem, these are precisely the 2-edge-connected graphs. However this is not sufficient for the biological problem, due to the additional constraints imposed by SNARE interaction rules. For example, if two edges of the graph have distinct target compartments, it must be they have distinct SNARE compositions to appropriately enforce those targets. While 2-edge-connectedness is not sufficient to guarantee this, 3-edge-connectedness is. This is because when edges of a strongly-connected 3-edge-connected graph are labeled by their cycle membership, no two edges belong to the same set of cycles.

In summary, the higher the edge-connectedness of a measured VTS graph, the more of the underlying biological constraints can be satisfied by appropriately placing SNAREs on each node and edge. If the data provided by cell biologists does not have a certain degree of connectedness, this implies that some biological data has been missed. Our goal here is to make this connection precise, by framing these constraints as a SAT instance.

It is perhaps counter-intuitive that, beyond a certain degree of edge-connectedness, any VTS graph can be implemented as a biological system. This is because we have not placed any constraints on the number of distinct SNAREs and other trafficking molecules a cell contains. In practice there is an upper bound on the number of molecules a cell can genetically encode. This question (``What is the maximum edge-connectedness possible given a certain number of SNAREs?'') is complementary to the question we have asked here (``What is the minimum edge-connectedness possible given any number of SNARES operating under certain physical constraints?'').

### 2019-09-22 papers/journal/intro.tex
Mukund's push

Nonetheless, completing the VTS for different organisms is very useful. 
%
The malfunction of the VTS causes many diseases.
%
The knowledge of the complete network would help identify the root causes of
the diseases~\cite{bexiga2013human,gissen2007cargos}. 
%
At a more basic level, having complete pictures of the VTS for various organisms and various cell-types would allow comparative studies, and therefore would allow the deduction of modes of evolution of the form of the vesicle traffic network. 
%
knowing the vesicle traffic network would also allow us to decipher features such as parts of the traffic system that are unchanging and therefore likely to be its core, and the more plastic parts~\cite{barlow2017seeing}.

@@

In previous work, a sampling approach was used. 
%
The study used a rule-based approach~\cite{mani2016stacking}.
Given an initial condition of the cell and rules for molecular traffic,
the approach updates the state of each compartment in the cell until the cell reaches a steady-state. 
%
With this study, statistical claims can be made, but improbable albeit biologically possible solutions could have been missed.
%
Also, there are many ways in which the dynamics of vesicle transport can be played out, while a single one was explored in this work. 
%
It is possible that a different scheme for the dynamics could lead to a different set of solutions.

>>>>

Nonetheless, completing the VTS for different organisms is very useful. 
%
The malfunction of the VTS causes many diseases.
%
The knowledge of the complete network would help identify the root causes of
the diseases~\cite{bexiga2013human,gissen2007cargos}. 
%
At a more basic level, having complete pictures of the VTS for various organisms and various cell-types would allow comparative studies, and therefore would allow the deduction of modes of evolution of the form of the vesicle traffic network. 
%
Knowing the vesicle traffic network would also allow us to decipher features such as parts of the traffic system that are unchanging and therefore likely to be its core, and the more plastic parts~\cite{barlow2017seeing}.

@@

In previous work, a sampling approach was used. 
%
The study used a rule-based specification~\cite{mani2016stacking}.
Given an initial condition of the cell and rules for molecular traffic,
that approach updates the state of each compartment in the cell until the cell reaches a steady-state. 
%
With this previous study, statistical claims can be made, but improbable albeit biologically possible solutions could have been missed.
%
Also, this previous study did not consider specific mechanisms of SNARE regulation. 
%
It is possible that a different scheme for the SNARE regulation could lead to a different set of solutions.
